---
title: "Lab 2"
author: "Christian Uriostegui"
date: "2022-12-08"
output: html_document
---

# Objective

My objective is to upload a movie survey into a PostgresSQL database, and connect it to R studio.

# Load Library

```{r}
library(RPostgres)
library(DBI)
library(dplyr)
```

# Establish connection between Postgres and R

Now that I have the neccessary libraries, I connect to my Postgres account by using my credentials.

```{r}
# Plug in credentials
cred<-dbConnect(RPostgres::Postgres(), dbname = "postgres", host="localhost", port="5432", user="postgres", password="Kristian7389!")   
```

Once connected, I then use the function below to ensure I am properly connected and verifying that my tables of interest are there.

```{r}
dbListTables(cred)
```

# Create Dataframe

After my object is created, I then transform it into a database while also selecting columns of interest.

```{r}
df <- dbGetQuery(cred, "SELECT gender,score, movie  FROM survey")
```

# Look at Review By Gender

```{r}
df %>%
  group_by(gender,movie) %>%
  summarise(average = mean(score)) %>%
  arrange(desc(average))
```

We can see that in my survey, the best reviewed movie was Get Out by both male and females.

# Disconnect

Once complete, we can now disconnect from the database.

```{r}
dbDisconnect(cred)
```
