---
title: "Assignment 1"
author: "Christian Uriostegui"
date: "2022-09-05"
output: 
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 150
---

# Introduction

The dataset to be examined contains political contributions from team owners and commissioners in major sports - which include the NFL, NBA, WNBA, NHL, MLB, and NASCAR. Column information includes specifics on the amount contributed and the recipients of these donations.

In my study I will examine how much contributions commissioners make towards political campaigns.

# Load Library

```{r message=FALSE}
library(dplyr)
library(tidyverse)
library(stringr)
```

# Data upload

```{r}
data <- 'https://raw.githubusercontent.com/curiostegui/CUNY-SPS/main/Data%20607/Assignment%20-%20%20Loading%20Data%20into%20a%20Data%20Frame/sports-political-donations.csv'
donation <- read.csv(file = data, header = TRUE, sep = ",")
head(donation)
```

# Data Transformation

Using the dplyr package I created a subset of the original donation set, filtering specifically for commissioners. In order to get the total sum of donations, I also have to change the `amount` column into a numeric column.

```{r}
commissioner <- donation %>%
  filter(Team == "Commissioner")
glimpse(commissioner)
```

```{r}
commissioner$Amount = as.numeric(gsub("[\\$,]", "",commissioner$Amount))
```

# Data Exploration

```{r}
commissioner_total <- commissioner %>%
  group_by(Owner, League) %>%
  summarise(Total = sum(Amount)) %>%
  arrange(desc(Total))
commissioner_total
```

```{r}
ggplot(commissioner_total, aes(x=Owner,y=Total)) +
       geom_col()
```

# Conclusion

We can see in the table and our visualization that commissioner Adam Silver, from the NBA, spends the most in polticial contribution with $22400. The next highest spender is Roger Goodell with $21320 in total spending from 2016-2020.

In a future study, I would want to explore how much commissioner spend on contributions relative to other non-commissioner owners.

