---
title: "Data 607 Assignment Week 3"
author: "Christian Uriostegui"
date: "2022-09-18"
output: html_document
---

```{r load data}
data <- 'https://raw.githubusercontent.com/curiostegui/CUNY-SPS/main/Data%20607/Assignment%20Week%203/majors-list.csv'
majors <- read.csv(file = data, header = TRUE, sep = ",")
```

```{r package}
library(stringr)
```

## Exercise 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset, provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
grep(pattern = 'data|science', majors$Major, value = TRUE, ignore.case = TRUE)
```

## Exercise 2

_Write_ _code_ _that_ _transforms_ _the_ _data_ _below_

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

_Into_ _a_ _format_ _like_ _this_:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r}
food <- list(c("bell pepper","bilberry","blackberry","blood orange", "blueberry","cantaloupe","chili pepper","cloudberry" ,"elderberry","lime","lychee","mulberry","olive","salal berry"))
```

Created a list and then used the cat and paste function to transform list

```{r}
cat(paste(food), collapse = ",")
```


## Exercise 3
Describe, in words, what these expressions will match:

***(.)\1\1***

What is returned will start with any character and be followed be with \1\1. For ex: 3\1\1, c\1\1


***"(.)(.)\\2\\1"***

Will return string any two characters, that follows up with a reversal of two characters. For ex: 1221, 4554, Chhc


***(..)\1***

Will return any two sets of character that is then followed by \1. For ex: 


***"(.).\\1.\\1"***

Returns strings that have a chracter that repeats in the middle and last part of the string


***"(.)(.)(.).*\\3\\2\\1"***

Will return strings where the first 3 characters can be any character and is then reversed in order

## Exercise #4

Construct regular expressions to match words that:

1. Start and end with the same character

```{r}
exer1 = 'creiucdiu'
  
exer_str <- str_extract_all(exer1, "(.).+\\1")
print(exer_str)
```

2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
exer2 = 'cacehjhhcadejnk'

exer2_str <- str_extract_all(exer2, "(..).*\\1")
print(exer2_str)
```

3. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
exer3 = 'banana'
  
exer3_str <- str_extract_all(exer3, "(.).*\\1.*\\1")
print(exer3_str)
```
