---
title: "API"
author: "Christian Uriostegui"
date: "2022-10-28"
output: html_document
---

# Assignment

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis

You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and 
transform it into an R DataFrame

# Load library

```{r library}
library(httr)
library(jsonlite)
library(tidyr)
library(dplyr)
```

# Connect to NYT API

I connected to the NYT Books API using the GET function. 

```{r}
res = GET("https://api.nytimes.com/svc/books/v3/lists/full-overview.json?api-key=wAWL9GvlAEQXqPqHPdP6g5NKswYDCt26")
```

I successfully connected based on the 200 status returned.

```{r}
res
```

# Transform Data to Dataframe

In this step, I turn the data into a Dataframe.

```{r}
#  the content here is turned from its raw form into a JSON style format and then turned to dataframe
data = fromJSON(rawToChar(res$content)) %>%
  data.frame()
```

# Further Cleaning

When looking at the dataframe I noticed that in the last column there appears to be a nested dataframe. Some of the more interesting information such as book title and author is in there.

```{r}
data %>%
  select(14:17) %>%
  as_tibble()
```

## Unnest Data

To access the information under the results.lists.books column, I use the function unnest.

```{r}
data <- unnest(data, results.lists.books)
```

```{r}
head(data)
```

## Select data of interest

Lastly, I selected the columns of interest such as title and author. If I were to do an analysis, I would work with the columns selected in this new set

```{r}
data <- data %>%
  select(c(35,36,40,38,20,26,34,12,5,7,4,13))
  as_tibble()
```

```{r}
data
```
