---
title: "Project 2 Lab 607"
author: "Christian Uriostegui"
date: "2022-10-09"
output: html_document
---

## Task

Using three datasets identified in the discussion board items or your own dataset, create a .CSV file that includes all of the information in the dataset; read the information from your CSV file into R and use tidyr and dplyr as needed to to tidy and transform data; perform analysis, and present conclusions.

## Introduction

Anthony Fontano is a music critic who creates YouTube content which detail his reviews of music projects from all genres. His review scores have come under attack from the artists he reviews - most recently by Yasiin Bey of Hip Hop group Black Star. In this project, I will be looking at the ratings of music projects and seeing if he has any preferences on project types (album,mixtape,track), genres, or if his tastes are influenced by popularity (spotify ranking).

## Dataset

I will be utilizing an Anthony Fontano album review dataset from Kaggle. Using a random sample of 20 rows, I created the projects dataset. For the 2nd and 3rd dataset, I created seperate files which list the genre and spotify rank of the artists from the projects dataset.

## Load library

```{r library, message = FALSE}
library(tidyr)
library(dplyr)
```

## Load datasets

```{r setup}
data <- 'https://raw.githubusercontent.com/curiostegui/CUNY-SPS/main/Data%20607/Project%202/sample_reviews.csv'
projects <- read.csv(file = data, header = TRUE, sep = ",")

data2 <- 'https://raw.githubusercontent.com/curiostegui/CUNY-SPS/main/Data%20607/Project%202/genre_artist.csv'
genre <- read.csv(file = data2, header = TRUE, sep = ",")

data3 <- 'https://raw.githubusercontent.com/curiostegui/CUNY-SPS/main/Data%20607/Project%202/rank_artists.csv'
rank <- read.csv(file = data3, header = TRUE, sep = ",")
```

## Tidy Data

### Projects dataset

Looking at the projects dataset, we see three problems: The wide format, Null values, and an extra column that needs to be removed. 
```{r}
as_tibble(projects)
```

Below I use the gather function to join and create new columns that seperate objects by the project type and the name. I also remove the extra column through the select function.

```{r}
projects <- gather(projects,"project_type","project_name",7:9)
```

```{r}
projects <- projects %>%
  filter(projects$project_name != "NA") %>%
  select('X','year','tracks','artist','rating','project_type','project_name')
```

### Genre dataset

Here I will turn the wide format into a tidy dataframe.

```{r}
as_tibble(genre[,1:9])
```

I utilize the gather and select function to tidy the data.

```{r}
genre <- genre %>%
  gather("artist","genre",2:21)
```

```{r}
genre <- select(genre,"artist","genre")
```

### Rank dataset


Like the previous dataset, I will also tidy the spotify rank dataframe.

```{r}
as_tibble(rank[,1:9])
```

I utilize the gather and select function to tidy the data.

```{r}
rank <- rank %>%
  gather("artist","rank",2:21)
```

```{r}
rank <- select(rank,"artist","rank")
```


### Fix typos

I fixed a typo I spotted in one of the artist names for both the rank and genre dataset.

```{r}
genre$artist[3] = "röyksopp"
rank$artist[3] = "röyksopp"
```



### Join

After tidying the data, I decided to join all the datasets so that I can conduct my analysis.

```{r}
joined <- left_join(genre,rank,by='artist')
```

I observed a period in between some of the artist names. This can pose a problem when joining with the projects set, since the artist names in that set are separated by a space.

```{r}
as_tibble(joined)
```

Below I remove the period in between some of the artist names.

```{r}
joined$artist[4] = "mac demarco"
joined$artist[8] = "erykah badu"
joined$artist[11] = "lil wayne"
joined$artist[12] = "trippie redd"
joined$artist[13] = "young thug"
joined$artist[14] = "blood orange"
joined$artist[15] = "doja cat"
joined$artist[16] = "tame impala"
joined$artist[17] = "summer walker"
joined$artist[18] = "young stoner life, young thug, gunna"
joined$artist[19] = "ibibio sound machine"
```

### Final dataset

```{r}
finalset <- left_join(projects,joined,by='artist')
```

```{r}
as_tibble(finalset)
```


## Analysis

In his top five highest reviews, we can observe that with the exception of Summer Walker, the other artists are not very popular, being that they don't appear in Spotify's top 500 streaming ranking. The genre of his top five reviews are rock,alternative,indie and R&B.

```{r}
as_tibble(finalset) %>% arrange(desc(rating))
```

The genre with the highest reviews is the rock genre. The lowest reviewed genre is EDM

```{r}
as_tibble(finalset) %>% group_by(genre) %>%
  summarise(mean_rating=mean(rating)) %>%
  arrange(desc(mean_rating))
```

The project type with the highest reviews is singles(tracks). Album projects had the lowest reviews.

```{r}
as_tibble(finalset) %>% group_by(project_type) %>%
  summarise(mean_rating=mean(rating)) %>%
  arrange(desc(mean_rating))
```

## Conclusion

We can observe that in **this** dataset, Anthony likes the rock genre; the artists that have the best reviews tend to not be on Spotify's top streaming charts, and his favorite project type are individual tracks. It is important to not that this was a very small subset of over thousands of album reviews he has done. 

Within this small dataset, we can see that there is an overwhelming number of mixtapes. We can also observe that his highest reviewed genre rock, only comprises 2 out of the 20 album reviews. The high reviews can skew our summary.


```{r}
finalset %>%
  count(project_type) 
```

```{r}
finalset %>%
  count(genre)
```

Because of the limitations listed above, we can not definitively state what his musical tastes are. If I had to conduct the study again, I would choose a larger dataset.
